\documentclass{article}
\usepackage[utf8]{inputenc}
\title{Dynamic Model in the Proximal tubule of Rat Kidney}
\author{Jing Yang}
\date{November 2019}

\usepackage{natbib}
\usepackage{graphicx}
\usepackage{amsmath}


\begin{document}

\maketitle

\begin{abstract}
We developed a dynamic model of a male rat proximal tubule in order to investigate results brought by changes on different lumen boundary conditions. We examined how long it will take to reach a steady state again from a sudden change. We also investigated whether oscillation on a single parameter will result in similar oscillations on other parameters. We determined that these oscillations will not have a significant passive effect on proximal tubule's re-absorption. We also noticed several irregular phenomenon such as phase difference and amplitude difference on the oscillation pattern of concentrations in different part along the tubule. And we tried to explain these response by using physiological knowledge.\\

\noindent \textbf{Keywords: }Proximal tubule $\cdot$ Oscillations $\cdot$ Concentrations $\cdot$ Dynamic model
\end{abstract}


\section{Model Formulation}
In proximal tubule, cells and tight junctions between cells form a single-layer cell wall between lumen and bath. We build our model based on the steady state model that Anita Layton published. The model consider fifteen kinds of solutes, including $\rm {Na}^{+}$,$\rm {K}^{+}$,$\rm {Cl}^{-}$,$\rm {HCO}_{3}^{-}$,$\rm H_{2}CO_{3}$,$\rm CO_{2}$,$\rm HPO_{4}^{2-}$,$\rm H_{2}PO_{4}^{-}$,urea,$\rm NH_{3}$,$\rm NH_{4}^{+}$,$\rm H^{+}$,\\$\rm HCO_{2}^{-}$,$\rm H_{2}CO_2$ and glucose. Model contains four compartments (Lumen, cell, lateral space, bath).Solutes are exchanged between different compartments through diffusion, transporters and coupled transporters. System equations are based on mass conservation and electronic neutrality. 


\subsection{Water Conservation}
Water conservation in cell and lateral space (denoted by subscripts ``C'' and ``P'' respectively) is given by:\\
\begin{equation}
\frac{\mathrm{d} V_{\mathrm{C}}}{\mathrm{d} t}=J_{v, \mathrm{LC}}+J_{v, \mathrm{BC}}+J_{v, \mathrm{PC}}
\end{equation}
\begin{equation}
\frac{\mathrm{d} V_{\mathrm{p}}}{\mathrm{d} t}=J_{v, \mathrm{LP}}+J_{v, \mathrm{BP}}+J_{v, \mathrm{CP}}
\end{equation}
where the subscripts ``L'' and ``B'' denote lumen and bath respectively, and ``v'' denotes volume of water. Water conservation in lumen is given by:
\begin{equation}
0=-\frac{\partial \mathrm{F}_{v}}{\partial x}+2 \pi r (J_{v,\mathrm{CL}}+J_{v,\mathrm{PL}})
\end{equation}
where $\mathrm{F}_{v}$ means water flow in lumen, $r$ means radius of lumen and partial derivative means along the lumen.

\subsection{Solute Conservation}
For non-reacting solute $k$, conservation equations are given by:
\begin{equation}
\frac{\mathrm{d} V_{\mathrm{C}} C_{k, \mathrm{C}}}{\mathrm{d} t}=J_{k, \mathrm{LC}}+J_{k, \mathrm{BC}}+J_{k, \mathrm{PC}}
\end{equation}
\begin{equation}
\frac{\mathrm{d} V_{\mathrm{P}} C_{k, \mathrm{P}}}{\mathrm{d} t}=J_{k, \mathrm{LP}}+J_{k, \mathrm{BP}}+J_{k, \mathrm{CP}}
\end{equation}
\begin{equation}
0=-\frac{\partial\left(\mathrm{F}_{v} C_{k,\mathrm{L}}\right)}{\partial x}+2 \pi r (J_{k,\mathrm{CL}}+J_{k,\mathrm{PL}})
\end{equation}
where $C_{k,i}$ denotes the concentration of solute $k$ in compartment $i$ and $J_{k,ij}$ denotes flux of solute $k$ from compartment $i$ to $j$.\\

For the reacting solutes, conservation is applied to the total buffers. Take the buffer $\mathrm{HPO}_{4}^{2-} and {H}_{2} \mathrm{PO}_{4}^{-}$ for instance:
\begin{equation}
\frac{\mathrm{d}}{\mathrm{d} t}\left(V_{i}\left(C_{\mathrm{HPO}_{4}^{2-}, i}+C_{\mathrm{H}_{2} \mathrm{PO}_{4}^{-}, i}\right)\right)=\hat{J}_{\mathrm{HPO}_{4}^{2-}, i}+\hat{J}_{\mathrm{H}_{2} \mathrm{PO}_{4}^{-}, i}
\end{equation}
where $i$ corresponding to C or P. $\hat{J}_{k, i}$ denotes the net flux of solute k into the compartment i.
\begin{equation}
\frac{\mathrm{d}}{\mathrm{d} t}\left(F_{v}\left(C_{\mathrm{HPO}_{4}^{2-}, L}+C_{\mathrm{H}_{2} \mathrm{PO}_{4}^{-}, L}\right)\right)=\hat{J}_{\mathrm{HPO}_{4}^{2-}, L}+\hat{J}_{\mathrm{H}_{2} \mathrm{PO}_{4}^{-}, L}
\end{equation}
The other buffer pairs $\left(\mathrm{CO}_{2},\mathrm{HCO}_{3}^{-}, \mathrm{H}_{2} \mathrm{CO}_{3}\right),\left(\mathrm{HPO}_{4}^{2-}, \mathrm{H}_{2} \mathrm{PO}_{4}^{-}\right),\left(\mathrm{NH}_{3}, \mathrm{NH}_{4}^{+}\right)$ all have above equations.
The buffer pairs are in equilibrium:
\begin{equation}
\mathrm{pH}=\mathrm{pK}_{A}-\log \frac{C_{A, i}}{C_{B, i}}
\end{equation}
where the buffer pairs (A,B) are $\left(\mathrm{HCO}_{3}^{-}, \mathrm{H}_{2} \mathrm{CO}_{3}\right),\left(\mathrm{HPO}_{4}^{2-}, \mathrm{H}_{2} \mathrm{PO}_{4}^{-}\right),\left(\mathrm{NH}_{3}, \mathrm{NH}_{4}^{+}\right),\left(\mathrm{HCO}_{2}^{-}, \mathrm{H}_{2} \mathrm{CO}_{2}\right)$.
The conservation of hydrogen is given by:
\begin{equation}
\begin{aligned}
\frac{\mathrm{d}}{\mathrm{d} t}\left(V_{i} (C_{\mathrm{H}^{+}}+C_{\mathrm{NH}_{4}^{+}}-C_{\mathrm{HCO_{3}^{-}}}-C_{\mathrm{HPO_{4}^{2-}}}-C_{\mathrm{HCO_{2}^{-}}}) \right)\\=\sum (\hat{J}_{H^{+}, i}+\hat{J}_{\mathrm{NH}_{4}^{+}, i}-\hat{J}_{ \mathrm{HCO_{3}^{-}, i}}-\hat{J}_{ \mathrm{HPO_{4}^{2-}, i}}-\hat{J}_{ \mathrm{HCO_{2}^{-}, i}})
\end{aligned}
\end{equation}

\subsection{Calculation}
We discrete time-dependent term by Back-Euler Method. For the partial differential term, we simply regard $F_{v}$ does not change too much along one cell's length along the tubule. For instance, solute conservation equation in lumen can be discrete as following:
For non-reacting solute $k$, conservation equations are given by:
\begin{equation}
\frac{V_{\mathrm{C}}^{n+1} C_{k, \mathrm{C}}^{n+1}-V_{\mathrm{C}}^{n} C_{k, \mathrm{C}}^{n}}{\Delta t}=J_{k, \mathrm{LC}}^{n+1}+J_{k, \mathrm{BC}}^{n+1}+J_{k, \mathrm{PC}}^{n+1}
\end{equation}
where ${\Delta t}$ means time step length, superscript $n$ means the n-th time step.

\section{Numerical Experiment}
After finishing the building of dynamic model, we can do various numerical experiments to observe changes along the time. In general, we did four kinds of experiments on boundary conditions separately :
\begin{itemize}
	\item Increase a certain solute concentration level
	\item Add various frequency and amplitude oscillations on SNGFR 
	\item Add various frequency and amplitude oscillations a certain solute concentration
	\item Add oscillations on solute concentration and SNGFR simultaneously
\end{itemize}
And we found several interesting results and tried to explain them by physiological knowledge.

\subsection{Experiment 1}
We increase glucose and urea concentration in lumen boundary condition separately. The proximal tubule plays an important role in reabsorption. Glucose are transported through transporter SGLT2 between lumen and cell,through transporter GLUT2 between cell and bath. Urea is a kind of inorganic chemicals with small molecule. It is transported by diffusion. Glucose and urea are both chemicals with no valence, meanwhile their concentrations are close in lumen. We calculate the convergence process to exam their difference in reabsorption. Since the transporter will accelerate the process


\section{Numerical Experiment}
``I always thought something was fundamentally wrong with the universe'' \citep{adams1995hitchhiker}

\bibliographystyle{plain}
\bibliography{references}
\end{document}
